Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-24 WSSSPE1_report/WSSSPE1_summary_report.tex
Update WSSSPE1_summary_report.tex
This commit includes suggested updates for the abstract by Anne C. Elster
Further discussion on software licenses and some suggested updates of the 
conclusions will follow in subsequent commits.

\author{Daniel~S.~Katz$^{(1)}$,
Sou-Cheng~T.~Choi$^{(2)}$,
Hilmar~Lapp$^{(3)}$,
Ketan~Maheshwari$^{(4)}$,
Frank~L\"{o}ffler$^{(5)}$,
Matthew~Turk$^{(6)}$,
Marcus~D.~Hanwell$^{(7)}$,
Nancy~Wilkins-Diehr$^{(8)}$,
James~Hetherington$^{(9)}$,
James~Howison$^{(10)}$,
Shel~Swenson$^{(11)}$,
Gabrielle~D.~Allen$^{(12)}$,
Bruce~Berriman$^{(13)}$
Colin~Venters$^{(14)}$
}

\thanks{{}$^{(1)}$ National Science Foundation, Arlington, VA, USA; Computation Institute, University of Chicago \& Argonne National Laboratory, Chicago, IL, USA; \url{d.katz@ieee.org}}
%
\thanks{{}$^{(2)}$ NORC at the University of Chicago and   Illinois Institute of Technology, Chicago, IL, USA; \url{sctchoi@uchicago.edu}}
%
\thanks{{}$^{(3)}$ National Evolutionary Synthesis Center (NESCent),
  Durham, NC, USA; \url{hlapp@nescent.org}}
%
\thanks{{}$^{(4)}$ Argonne National Laboratory, Chicago, IL, USA; \url{ketan@anl.gov}}
%
\thanks{{}$^{(5)}$ Center for Computation and Technology, Louisiana State University, Baton Rouge, LA, USA; \url{knarf@cct.lsu.edu}}
%
\thanks{{}$^{(6)}$ Columbia Astrophysics Laboratory, Columbia University, New
York, NY, USA; \url{matthewturk@gmail.com}}
%
\thanks{{}$^{(7)}$ Scientific Computing Group, Kitware, Inc.,  Clifton Park, NY, USA; \url{marcus.hanwell@kitware.com}}
%
\thanks{{}$^{(8)}$ San Diego Supercomputer Center, University of California San Diego, La Jolla, CA, USA, \url{wilkinsn@sdsc.edu}}
%
\thanks{{}$^{(9)}$ Research Software Development Team, University College London, London, UK; \url {j.hetherington@ucl.ac.uk}}
%
\thanks{{}$^{(10)}$ University of Texas at Austin, Austin, TX, USA; \url{jhowison@ischool.utexas.edu}}
%
\thanks{{}$^{(11)}$ University of Illinois, Champaign, IL, USA; \url{gdallen@illinois.edu}}
%
\thanks{{}$^{(12)}$ University of Southern California, Los Angeles, CA, USA; \url{mdswenso@usc.edu}}
%
\thanks{{}$^{(11)}$ Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA, USA; \url{gbb@ipac.caltech.edu}}
%
\thanks{{}$^{(14)}$ University of Huddersfield, Huddersfield, West Yorkshire, UK; \url{C.Venters@hud.ac.uk}}

\begin{abstract} %\note{Lead: Sou-Cheng T. Choi}
This document summarizes the First Workshop on Sustainable Software
for Science: Practice and Experiences (WSSSPE1). The content is based
on a set of collaborative notes taken during the workshop, including
discussion about the keynote presentations, three themed panels that
discussed papers accepted at the workshop, other discussion,
cross-cutting issues, use cases, and conclusions. It joins a nascent
literature seeking to understand what drives software work in science
and how the reward systems of science function, thereby shaping the
types of software projects undertaken. These incentives can determine the extent to which
developers are motivated to build software for the long-term, for the
use of others, and whether to work collaboratively or separately.
Unique perspectives were captured about issues such as comprehensive
documentation, development and deployment practices, software
licenses and career paths for developers. Attribution systems that capture
evidence of software contribution and impact were also discussed. These include
mechanisms such as Digital Object Identifiers, publication
of ``software papers'' and the use of online systems, for example source code repositories
like GitHub. For truly sustainable software, there should be no
endpoint, as the software products continue to be used beyond a single
institution, grant, and developer or development team.

@@

Additional volunteers:
\begin{itemize}
% \item Chris Mattmann $\langle$\url{chris.a.mattmann@jpl.nasa.gov}$\rangle$
\item * Ketan Maheshwari $\langle$\url{ketancmaheshwari@gmail.com}$\rangle$
\item Marlon Pierce $\langle$\url{marpierc@iu.edu}$\rangle$
\item * Colin Venters $\langle$\url{C.Venters@hud.ac.uk}$\rangle$
\item Suresh Marru $\langle$\url{smarru@iu.edu}$\rangle$
% \item Lynn Zentner $\langle$\url{lzentner@purdue.edu}$\rangle$
\item Anne Elster $\langle$\url{anne.elster@gmail.com}$\rangle$
\item * Shel Swenson $\langle$\url{mdswenso@usc.edu}$\rangle$
\item * Andy Ray Terrel $\langle$\url{andy.terrel@gmail.com}$\rangle$
% \item Abani Patra $\langle$\url{abani.patra@gmail.com}$\rangle$
\item * Nancy Wilkins-Diehr $\langle$\url{wilkinsn@sdsc.edu}$\rangle$
% \item James Spencer $\langle$\url{j.spencer@imperial.ac.uk}$\rangle$
\item * Frank L\"{o}ffler $\langle$\url{knarf@cct.lsu.edu}$\rangle$
\item * James Hetherington $\langle$\url{j.hetherington@ucl.ac.uk}$\rangle$
\item * Sou-Cheng (Terrya) Choi $\langle$\url{sctchoi@uchicago.edu}$\rangle$
\item * James Howison $\langle$\url{james@howison.name}$\rangle$
\item * Bruce Berriman $\langle$\url{gbb@ipac.caltech.edu}$\rangle$
\item * Hilmar Lapp $\langle$\url{hlapp@nescent.org}$\rangle$
\item * Marcus D. Hanwell $\langle$\url{marcus.hanwell@kitware.com}$\rangle$
\item Lucas Nussbaum $\langle$\url{lucas.nussbaum@loria.fr}$\rangle$
\item * Matthew Turk $\langle$\url{matthewturk@gmail.com}$\rangle$
\end{itemize}

@@

>>>>

\author{Daniel~S.~Katz$^{(1)}$,
Sou-Cheng~T.~Choi$^{(2)}$,
Hilmar~Lapp$^{(3)}$,
Ketan~Maheshwari$^{(4)}$,
Frank~L\"{o}ffler$^{(5)}$,
Matthew~Turk$^{(6)}$,
Marcus~D.~Hanwell$^{(7)}$,
Nancy~Wilkins-Diehr$^{(8)}$,
James~Hetherington$^{(9)}$,
James~Howison$^{(10)}$,
Shel~Swenson$^{(11)}$,
Gabrielle~D.~Allen$^{(12)}$,
Anne~C.~Elster$^{(13)}$,
Bruce~Berriman$^{(14)}$
Colin~Venters$^{(15)}$
}

\thanks{{}$^{(1)}$ National Science Foundation, Arlington, VA, USA; Computation Institute, University of Chicago \& Argonne National Laboratory, Chicago, IL, USA; \url{d.katz@ieee.org}}
%
\thanks{{}$^{(2)}$ NORC at the University of Chicago and   Illinois Institute of Technology, Chicago, IL, USA; \url{sctchoi@uchicago.edu}}
%
\thanks{{}$^{(3)}$ National Evolutionary Synthesis Center (NESCent),
  Durham, NC, USA; \url{hlapp@nescent.org}}
%
\thanks{{}$^{(4)}$ Argonne National Laboratory, Chicago, IL, USA; \url{ketan@anl.gov}}
%
\thanks{{}$^{(5)}$ Center for Computation and Technology, Louisiana State University, Baton Rouge, LA, USA; \url{knarf@cct.lsu.edu}}
%
\thanks{{}$^{(6)}$ Columbia Astrophysics Laboratory, Columbia University, New
York, NY, USA; \url{matthewturk@gmail.com}}
%
\thanks{{}$^{(7)}$ Scientific Computing Group, Kitware, Inc.,  Clifton Park, NY, USA; \url{marcus.hanwell@kitware.com}}
%
\thanks{{}$^{(8)}$ San Diego Supercomputer Center, University of California San Diego, La Jolla, CA, USA, \url{wilkinsn@sdsc.edu}}
%
\thanks{{}$^{(9)}$ Research Software Development Team, University College London, London, UK; \url {j.hetherington@ucl.ac.uk}}
%
\thanks{{}$^{(10)}$ University of Texas at Austin, Austin, TX, USA; \url{jhowison@ischool.utexas.edu}}
%
\thanks{{}$^{(11)}$ University of Illinois, Champaign, IL, USA; \url{gdallen@illinois.edu}}
%
\thanks{{}$^{(12)}$ University of Southern California, Los Angeles, CA, USA; \url{mdswenso@usc.edu}}
%
\thanks{{}$^{(13)}$ Norwegian University of Science and Technology, Trondheim, Norway; \url{elster@ntnu.no}}
%
\thanks{{}$^{(14)}$ Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA, USA; \url{gbb@ipac.caltech.edu}}
%
\thanks{{}$^{(15)}$ University of Huddersfield, Huddersfield, West Yorkshire, UK; \url{C.Venters@hud.ac.uk}}

\begin{abstract} %\note{Lead: Sou-Cheng T. Choi}
% Anne: Added a lead-in motivator to address the submission as a Experience paper
% and tried to improve writing... ( I am assumeing no abstract word count limit...)
Challenges related to the development, deployment, and maintenance of 
reusable software for science is becoming an increasing concern. 
There are now many scientists whose research increasingly depends on 
the quality and availability of software they can use and build on. 
At the same time, software specialists have more than enough to do 
without having to "reinvent the wheel". To highlight some of these 
issues and share experiences, The First Workshop on Sustainable 
Software for Science: Practice and Experiences (WSSSPE1) was held 
in November 2013 at SC13.
The workshop featured keynote presentations and based its discussion 
them, as well as on a large number (54) of solicited extended abstracts 
that were grouped into three themes presented via panels that encouraged 
audience participation. A set of collaborative notes was taken during the workshop.

This document 
% article? 
summarizes 
the issues and shared experiences discussed atthe workshop, including
cross-cutting issues and use cases. It joins a nascent
literature seeking to understand what drives software work in science,
and how it is impacted by the reward systems of science. 
These incentives can determine the extent to which
developers are motivated to build software for the long-term, for the
use of others, and whether to work collaboratively or separately.

Unique perspectives were captured about issues such as comprehensive
documentation, development and deployment practices, software
licenses and career paths for developers. Attribution systems that capture
evidence of software contribution and impact were also discussed. These include
mechanisms such as Digital Object Identifiers, publication
of ``software papers'' and the use of online systems, for example source code repositories
like GitHub. For truly sustainable software, there should be no
endpoint, as the software products continue to be used beyond a single
institution, grant, and developer or development team.

% Original:
% This document summarizes the First Workshop on Sustainable Software
% for Science: Practice and Experiences (WSSSPE1). The content is based
% on a set of collaborative notes taken during the workshop, including
% discussion about the keynote presentations, three themed panels that
% discussed papers accepted at the workshop, other discussion,
% cross-cutting issues, use cases, and conclusions. It joins a nascent
% literature seeking to understand what drives software work in science
% and how the reward systems of science function, thereby shaping the
% types of software projects undertaken. These incentives can determine the extent to which
% developers are motivated to build software for the long-term, for the
% use of others, and whether to work collaboratively or separately.
% Unique perspectives were captured about issues such as comprehensive
% documentation, development and deployment practices, software
% licenses and career paths for developers. Attribution systems that capture
% evidence of software contribution and impact were also discussed. These include
% mechanisms such as Digital Object Identifiers, publication
% of ``software papers'' and the use of online systems, for example source code repositories
% like GitHub. For truly sustainable software, there should be no
% endpoint, as the software products continue to be used beyond a single
% institution, grant, and developer or development team.

@@

Additional volunteers:
\begin{itemize}
% \item Chris Mattmann $\langle$\url{chris.a.mattmann@jpl.nasa.gov}$\rangle$
\item * Ketan Maheshwari $\langle$\url{ketancmaheshwari@gmail.com}$\rangle$
\item Marlon Pierce $\langle$\url{marpierc@iu.edu}$\rangle$
\item * Colin Venters $\langle$\url{C.Venters@hud.ac.uk}$\rangle$
\item Suresh Marru $\langle$\url{smarru@iu.edu}$\rangle$
% \item Lynn Zentner $\langle$\url{lzentner@purdue.edu}$\rangle$
\item Anne C. Elster $\langle$\url{elster@ntnu.no}$\rangle$
\item * Shel Swenson $\langle$\url{mdswenso@usc.edu}$\rangle$
\item * Andy Ray Terrel $\langle$\url{andy.terrel@gmail.com}$\rangle$
% \item Abani Patra $\langle$\url{abani.patra@gmail.com}$\rangle$
\item * Nancy Wilkins-Diehr $\langle$\url{wilkinsn@sdsc.edu}$\rangle$
% \item James Spencer $\langle$\url{j.spencer@imperial.ac.uk}$\rangle$
\item * Frank L\"{o}ffler $\langle$\url{knarf@cct.lsu.edu}$\rangle$
\item * James Hetherington $\langle$\url{j.hetherington@ucl.ac.uk}$\rangle$
\item * Sou-Cheng (Terrya) Choi $\langle$\url{sctchoi@uchicago.edu}$\rangle$
\item * James Howison $\langle$\url{james@howison.name}$\rangle$
\item * Bruce Berriman $\langle$\url{gbb@ipac.caltech.edu}$\rangle$
\item * Hilmar Lapp $\langle$\url{hlapp@nescent.org}$\rangle$
\item * Marcus D. Hanwell $\langle$\url{marcus.hanwell@kitware.com}$\rangle$
\item Lucas Nussbaum $\langle$\url{lucas.nussbaum@loria.fr}$\rangle$
\item * Matthew Turk $\langle$\url{matthewturk@gmail.com}$\rangle$
\end{itemize}

@@

%Anne did not attend the whole workshop as she was busy as a SC14 Deouty Panels chair, but definately was there :-)
Anne C. Elster
